An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-10 scagnostics.Rmd
Removing unnecessary plots.

```{r neutronStarEq,  results="hide", fig.height=3, fig.cap="Final view when optimising the splines2d index for the neutron star merger dataset, showing the functional dependence between the mass, radius and lambda parameter of one of the neutron stars."}
set.seed(2018)
nsD <- read_csv("data/samples.csv") 
nsM <- nsD %>%
  select(-m2) %>% # remove m2 to avoid finding well known correlation
  rescale() # rescaling, this returns a matrix
if(!file.exists("cache/neutronStarEq.rda")){
  nseTour <- save_history(nsM, guided_tour(splineIndex()))
  nseTourFull <- as.list(interpolate(nseTour))
  save(nseTour, nseTourFull, file = "cache/neutronStarEq.rda")
} else {
  load("cache/neutronStarEq.rda")
}
iLast <- length(nseTourFull)
fProj <- nseTourFull[[iLast]]
dProj <- as.tibble(nsM %*% fProj)
ggplot(dProj, aes(V1,V2)) + geom_point() + theme(aspect.ratio=1)
```

@@

The resulting projection clearly shows how the data selects a region that is well defined as a combination of several parameters. From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection. Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{nsePlotOrig} (right), which we compare to the next closest bivariate plot shown on the left. It is clear that in this example additional information is found when considering general projections rather than focusing only on bivariate relations.

```{r nsePlotOrig}
p1 <- ggplot(nsD, aes(R1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
p3 <- ggplot(nsD, aes(R1/m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
grid.arrange(p1, p2, p3, ncol=3) 
```

@@

```{r bbhRes, fig.cap="Approximating the final view found in the TIC guided tour optimisation by simple parameter combinations."}
ggplot(as.tibble(bbhM), aes(time, ra+dec)) + geom_point()
ggplot(bbhD, aes(time, ra+pi*dec)) + geom_point()
```

>>>>

```{r neutronStarEq,  results="hide"}
set.seed(2018)
nsD <- read_csv("data/samples.csv") 
nsM <- nsD %>%
  select(-m2) %>% # remove m2 to avoid finding well known correlation
  rescale() # rescaling, this returns a matrix
if(!file.exists("cache/neutronStarEq.rda")){
  nseTour <- save_history(nsM, guided_tour(splineIndex()))
  nseTourFull <- as.list(interpolate(nseTour))
  save(nseTour, nseTourFull, file = "cache/neutronStarEq.rda")
} else {
  load("cache/neutronStarEq.rda")
}
iLast <- length(nseTourFull)
fProj <- nseTourFull[[iLast]]
dProj <- as.tibble(nsM %*% fProj)
```

@@

The resulting projection clearly shows how the data selects a region that is well defined as a combination of several parameters. From the matrix representation we read off that parameters m1, L1 and R1 are important in this projection. Using the original (non-rescaled) data and taking into account the difference in scale between m1 and R1 we can reproduce a similar picture shown in Fig \ref{fig:nsePlotOrig} (right), which we compare to the next closest bivariate plot shown on the left. It is clear that in this example additional information is found when considering general projections rather than focusing only on bivariate relations.

```{r nsePlotOrig, fig.cap="Comparison of guided tour final view (left), approximation based on original parameters (middle) and expected relation based on analysis setup (right)."}
p1 <- ggplot(dProj, aes(V1,V2)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
p2 <- ggplot(nsD, aes(R1-2*pi*m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
p3 <- ggplot(nsD, aes(R1/m1, L1)) + geom_point(alpha = 0.05) + theme(aspect.ratio=1)
grid.arrange(p1, p2, p3, ncol=3) 
```

@@
